### 2017-05-09 introduction.tex
Add files via upload

\noindent where $Q^*$ is net radiation at the surface, $Q_F$, $Q_H$, and $Q_E$ are fluxes of anthropogenic, sensible, and latent heat, $\Delta Q_A$ is net advection of surface heat to or from the volume of air above the surface, and $\Delta Q_G$ is net storage of heat by the substrate below. A change in T\textsubscript{surf} is translated to change in T\textsubscript{air} as the surface exchanges energy with the atmosphere above it, primarily via convection and radiation. As such, T\textsubscript{surf} and T\textsubscript{are} are inextricably linked, particularly near to the ground, and a city's altered surface characteristics and energy balance manifest in modifications to both surface and air temperature regimes.

@@

The notion that urban modifications to the environment produce elevated T\textsubscript{air} and T\textsubscript{surf} is not new, with the first formal studies of the urban effect on T\textsubscript{air} dating back to \citet{Howard1833}'s identification of "artificial warming" and "urban contamination" in his characterization of spatial patterns of T\textsubscript{air} in early 19\textsuperscript{th} century London. Conceptually, contemporary studies of urban climate do not stray far from \citet{Howard1833}, by seeking to isolate and quantify the urban signal in measured or modeled assessment of a given meteorological variable. However, as \citet{Howard1833} identified nearly two centuries ago, the sheer number of factors influencing direct measurement of climates makes experimental control nearly impossible in observational study. In any given observation of climate, myriad urban and non urban signals make it difficult to reliably differentiate urban influenced measures from those free of urban influence, degrading interstudy comparison. Even in the modeling domain, where experimental control is possible, the inherent complexity in simulating flows of energy, momentum, and mass at the scales necessary to represent physical phenomena over realistic urban environments is, as of yet, not feasible. Notwithstanding implementation of accurate, physically based schemes to represent urban climates in regional or global climate models. Thus, accurate assessment of how a city affects the climate is a deceptively challenging task. To address this, \citet{Lowry1977} presents a framework to isolate the urban effect in observation of a given meteorological variable. In his framework - termed the "Lowry method" - measurement of a meteorological variable ($V_M$) is the sum of forcings from the background climate and synoptic conditions ($V_B$), local topography ($V_L$), and the urban landscape ($V_H$)

@@

Study of the UHI with respect to T\textsubscript{air} (aUHI) and T\textsubscript{surf} (sUHI) tacitly falls into such a framework, by replacing $V_{M, \ U} - V_{M, \ R}$ with $\Delta$T\textsubscript{U - R}. However, in spite of its conceptual simplicity through the lens of the Lowry method, meta-analysis of observational aUHI studies suggests that that simplicity may not translate well to real world analysis. In a points based assessment of 190 aUHI studies published between 1950 and 2007, \citet{Stewart2011} found that nearly 50\% of studies are "scientifically indefensible". In his analysis, each study was given a "passing" or "failing" grade and ranked based on methodological quality in the following criteria: conceptual model, operational definitions, instrument specification, site metadata, site representativeness, number of replicates, weather control, surface control, and synchronisity. Of the 190 surveyed studies, only 13\% were derived from field sites that were sufficiently representative of the local-scale environment or lacked the meta-data to make such a determination, leading to potential biases from micro-scale terrain. 89\% lacked meta-data altogether. These findings suggest a significant gap between conceptual simplicity and practical realities in UHI analysis.

@@

In contrast to urban T\textsubscript{air} and aUHI assessment, where measurement techniques have not undergone significant evolution in recent memory and methodological problems arise in large part from conceptual shortcomings and simple carelessness, study of the urban effect on land T\textsubscript{surf} is relatively new and significantly more complex. As such, methodological problems in sUHI study can arise from both conceptual flaws \textit{and} instrument biases. In spite of these difficulties, the introduction of new methods, instruments, and analytical techniques in TIR remote sensing of urban environments has lead to rapid proliferation in study of urban T\textsubscript{surf} and the sUHI - much of which is focused on remote sensing of upwelling TIR. In the decades preceding this thesis, a combination of factors has lead to an increased focus on remote sensed study of urban T\textsubscript{surf}:

\begin{itemize}
	\item The proliferation of satellite and aerial thermal infrared (TIR) remote sensors and "openness" in data access policies.
	\item Improvements in sensor spatial, spectral, and radiometric resolutions. 
	\item An increased focus on the importance of the surface in both determining and understanding key near-ground micro-meteorological phenomena.
\end{itemize}

@@

Traditional methods for urban T\textsubscript{surf} measurement are subject to a suite of geometric and temporal biases. Geometric biases are the result of urban modification of surface structure and thermal and radiative properties combined with the 2-dimensional, narrow field-of-view (FOV) viewing geometry of conventional remote sensors:

@@

\noindent---\textit{Undersampling of surface 3-dimensionality by narrow-FOV remote sensors}.

Relative to typical non-urban and rural terrain, urban development adds significant 3-dimensionality to the surface. This added complexity results in undersampling of the surface by narrow-FOV remote sensors. A narrow-FOV sensor, by sampling surface 3-dimensionality in 2-dimensions, "sees" a different assemblage of urban surfaces depending on its viewing direction, resulting in a geometric bias towards some arrangement of facets; i.e. viewing from the nadir, a narrow-FOV sensor sees the urban surface as an array of rooftops and roads, the same sensor viewing the surface from an oblique angle sees some array of walls, rooftops, and roads depending on its orientation. For a sensor in the nadir, increasing FOV introduces sampling of vertical features. To illustrate how sensor-surface geometry influences sampling of the urban surface by narrow- and wide-FOV sensors, Figure \ref{model} shows FOVs for the three sensor orientation/type combinations.The vast majority of urban thermal remote sensing is done via a narrow-FOV sensor in the nadir.

@@

\noindent---\textit{Directional dependence in remote sensed urban T\textsubscript{surf}}.

In addition to modifying how a remote sensor samples the surface, the convoluted 3-dimensional structure of the urban surface modifies sunlit/shading regimes, resulting in strong micro-scale spatiotemporal contrasts in urban T\textsubscript{surf}, which depend on surface-sun geometry. These microscale variations in urban T\textsubscript{surf} are often amplified by significant geometric and spatial contrasts in the thermal admittance of common urban materials, which determines the diurnal amplitude of T\textsubscript{surf} for a given facet. Thus, when viewed from a narrow-FOV remote sensor with some inherent geometric bias, urban T\textsubscript{surf} is directionally dependent - constituting an "effective anisotropy" of urban T\textsubscript{surf}, shown in Figure \ref{anisot}. The qualifier "effective" is used to differentiate directional contrasts in urban T\textsubscript{surf} that arise from a city's 3-dimensional surface structure from those resulting from the non-Lambertian nature of individual urban facets (e.g. walls, rooftops, or roads). The magnitude of urban effective anisotropy can reach up to 10 \si{\kelvin} and is highly dependent on surface-sensor-sun geometry, surface structure, and urban materials \citep{Krayenhoff2016, Voogt1997}.

@@

Temporal biases in thermal remote sensing of urban areas occur across multiple time scales including:

\noindent---\textit{ Contamination by turbulence forced, high frequency fluctuations in urban T\textsubscript{surf}}.

Using time-sequential thermography \citet{Christen2012} found that many common urban fabric types (e.g. rooftops, walls, roads, and vegetation) display large micro-scale (second to minute) fluctuations in T\textsubscript{surf}. The magnitude of these is inversely related to surface thermal admittance and is, as of writing, poorly understood. Most thermal remote sensors provide instantaneous, rather than temporally averaged, T\textsubscript{surf} and thus are subject to contamination by high-frequency fluctuations. These biases are difficult to estimate in urban environments, where a large variety of fabric materials can produce significant geometric and spatial contrasts in thermal admittance - and thus directional and spatial variations in the magnitude of microscale fluctuations depending on the facet material types viewed by the sensor.

\noindent---\textit{Discontinuity in satellite overpass cycles.}

Overpass cycles for polar orbiting sensors are discontinuous. These sensors provide data for the vast majority of urban T\textsubscript{surf} study. Assessment of land T\textsubscript{surf} either sacrifices spatial resolution for a daily repeat cycle - as is the case with MODIS  - or sacrifices temporal resolution for a higher spatial resolution - ASTER or Landsat.

\noindent---\textit{Clear-sky bias.}

@@

As is the case with geometric biases, the magnitude of these temporal biases in remote sensed urban T\textsubscript{surf} is difficult to quantify without long-term ground truthing campaigns or significant interpolation. As such, the true temporal and geometric nature of urban effect of T\textsubscript{surf} is presently unknown.

@@

\noindent In an attempt to answer these questions, I did the following:

\begin{itemize}
	\item Developed and evaluated a method to retrieve atmospherically correction hemispherical radiometric urban surface temperatures from time-continuous measurements of upwelling longwave radiation.
	\item Compared urban surface temperatures and surface urban heat island magnitudes retrieved using the method to common remote sensed representations of the urban surface.
	\item Derived an eight month climatology of hemispherical urban surface temperatures to observe seasonal and diurnal patterns of the surface urban heat island effect.
\end{itemize}

>>>>

\noindent where $Q^*$ is net radiation at the surface, $Q_F$, $Q_H$, and $Q_E$ are fluxes of anthropogenic, sensible, and latent heat, $\Delta Q_A$ is net advection of surface heat to or from the volume of air above the surface, and $\Delta Q_G$ is net storage of heat by the substrate below. A change in T\textsubscript{surf} is translated to change in T\textsubscript{air} as the surface exchanges energy with the atmosphere above it, primarily via convection and radiation. As such, T\textsubscript{surf} and T\textsubscript{air} are inextricably linked, particularly near to the ground, and a city's altered surface characteristics and energy balance manifest in modifications to both surface and air temperature regimes.

@@

The notion that urban modifications to the environment produce elevated T\textsubscript{air} and T\textsubscript{surf} is not new, with the first formal studies of the urban effect on T\textsubscript{air} dating back to \citet{Howard1833}'s identification of "artificial warming" and "urban contamination" in his characterization of spatial patterns of T\textsubscript{air} in early 19\textsuperscript{th} century London. Conceptually, contemporary studies of urban climate do not stray far from \citet{Howard1833}, by seeking to isolate and quantify the urban signal in measured or modeled assessment of a given meteorological variable. However, as \citet{Howard1833} identified nearly two centuries ago, the sheer number of factors influencing direct measurement of climates makes experimental control nearly impossible in observational study. In any given observation of climate, myriad urban and non urban signals make it difficult to reliably differentiate urban influenced measures from those free of urban influence, degrading interstudy comparison. Development of physically based numerical models to represent physical processes affected by urbanization, and which permit experimental control, have shown promise \citep{Krayenhoff2014,Gastellu-Etchegorry1996,Masson2000,Martilli2002} yet the inherent complexity in simulating flows of energy, momentum and mass at the scales necessary to represent the full range of physical processes over realistic environments remains an on-going challenge. Thus, accurate assessment of how a city affects the climate is a deceptively challenging task. To address this, \citet{Lowry1977} presents a framework to isolate the urban effect in observation of a given meteorological variable. In his framework - termed the "Lowry method" - measurement of a meteorological variable ($V_M$) is the sum of forcings from the background climate and synoptic conditions ($V_B$), local topography ($V_L$), and the urban landscape ($V_H$)

@@

Study of the UHI with respect to T\textsubscript{air} (aUHI) - typically referring to T\textsubscript{air} measured from approximately 2 \si{\meter} above ground level, yielding a "canopy layer" UHI (sUHI) - and T\textsubscript{surf} (sUHI) tacitly falls into such a framework, by replacing $V_{M, \ U} - V_{M, \ R}$ with $\Delta$T\textsubscript{U - R}. However, in spite of its conceptual simplicity through the lens of the Lowry method, meta-analysis of observational aUHI studies suggests that that simplicity may not translate well to real world analysis. In a points based assessment of 190 aUHI studies published between 1950 and 2007, \citet{Stewart2011} found that nearly 50\% of studies are "scientifically indefensible". In his analysis, each study was given a "passing" or "failing" grade and ranked based on methodological quality in the following criteria: conceptual model, operational definitions, instrument specification, site metadata, site representativeness, number of replicates, weather control, surface control, and synchronisity. Of the 190 surveyed studies, only 13\% were derived from field sites that were sufficiently representative of the local-scale environment or lacked the meta-data to make such a determination, leading to potential biases from micro-scale terrain. 89\% lacked meta-data altogether. These findings suggest a significant gap between conceptual simplicity and practical realities in UHI analysis.

@@

In contrast to urban T\textsubscript{air} and aUHI assessment, where measurement techniques have not undergone significant recent evolution and methodological problems arise in large part from conceptual shortcomings and simple carelessness, study of the urban effect on land T\textsubscript{surf} is relatively new and significantly more complex. As such, methodological problems in sUHI study can arise from both conceptual flaws \textit{and} instrument biases. In spite of these difficulties, the introduction of new methods, instruments, and analytical techniques in TIR remote sensing of urban environments has lead to rapid proliferation in study of urban T\textsubscript{surf} and the sUHI - much of which is focused on remote sensing of upwelling thermal infrared radiation (TIR). In the decades preceding this thesis, a combination of factors has lead to an increased focus on remote sensed study of urban T\textsubscript{surf}:

\begin{itemize}
	\item The proliferation of satellite and aerial TIR remote sensors and "openness" in data access policies.
	\item Improvements in sensor spatial, spectral, and radiometric resolutions. 
	\item An increased focus on the importance of the surface in both determining and understanding key near-ground micro-meteorological phenomena.
	\item Interest in T\textsubscript{surf} for human thermal comfort, energy conservation, and sUHI mitigation applications. 
\end{itemize}

@@

Traditional methods for urban T\textsubscript{surf} measurement are subject to a suite of geometric and temporal biases. Geometric biases are the result of urban modification of surface structure and thermal and radiative properties combined with the  narrow field-of-view (FOV) viewing geometry of conventional remote sensors:

@@

\subsection*{\textnormal{\textit{Undersampling of surface 3-dimensionality by narrow-FOV remote sensors}}}

Relative to typical non-urban and rural terrain, urban development adds significant 3-dimensionality to the surface. This added complexity results in undersampling of the complete urban surface by narrow-FOV remote sensors. A narrow-FOV sensor "sees" a different assemblage of urban surfaces depending on its viewing direction, resulting in a geometric bias towards some arrangement of facets; i.e. viewing from the nadir, a narrow-FOV sensor sees the urban surface as an array of rooftops and roads, the same sensor viewing the surface from an oblique angle sees some array of walls, rooftops, and roads depending on its orientation. For a sensor in the nadir, increasing FOV introduces sampling of vertical features. To illustrate how sensor-surface geometry influences sampling of the urban surface by narrow- and wide-FOV sensors, Figure \ref{model} shows FOVs for the three sensor orientation/type combinations.The vast majority of urban thermal remote sensing is done via a narrow-FOV sensor viewing from the nadir.

@@

\subsection*{\textnormal{\textit{Directional dependence in remote sensed urban T\textsubscript{surf}}}}

In addition to modifying how a remote sensor samples the surface, the convoluted 3-dimensional structure of the urban surface modifies the surface radiation budget, resulting in strong micro-scale spatiotemporal contrasts in urban T\textsubscript{surf}, which depend on surface-sun geometry. These microscale variations in urban T\textsubscript{surf} are often amplified by significant geometric and spatial contrasts in the thermal admittance of common urban materials, which determines the diurnal amplitude of T\textsubscript{surf} for a given facet. Thus, when viewed from a narrow-FOV remote sensor with some inherent geometric bias, urban T\textsubscript{surf} is directionally dependent - constituting an "effective anisotropy" of urban T\textsubscript{surf}, shown in Figure \ref{anisot}. The qualifier "effective" is used to differentiate directional contrasts in urban T\textsubscript{surf} that arise from a city's 3-dimensional surface structure from those resulting from the non-Lambertian nature of individual urban facets (e.g. walls, rooftops, or roads). The magnitude of urban effective anisotropy can reach up to 10 \si{\kelvin} and is highly dependent on surface-sensor-sun geometry, surface structure, and urban materials \citep{Krayenhoff2016, Voogt1997}.

@@

Temporal biases in thermal remote sensing of urban areas occur across multiple time scales including: 
\subsection*{\textnormal{\textit{Contamination by turbulence forced, high frequency fluctuations in urban T\textsubscript{surf}}}}

Using time-sequential thermography \citet{Christen2012} found that many common urban fabric types (e.g. rooftops, walls, roads, and vegetation) display large micro-scale (second to minute) fluctuations in T\textsubscript{surf}. The magnitude of these is inversely related to surface thermal admittance and is currently, poorly understood. Most thermal remote sensors provide instantaneous, rather than temporally averaged, T\textsubscript{surf} and thus are subject to contamination by high-frequency fluctuations. These biases are difficult to estimate in urban environments, where a large variety of fabric materials can produce significant geometric and spatial contrasts in thermal admittance --- and thus directional and spatial variations in the magnitude of microscale fluctuations depending on the facet material types viewed by the sensor.

\subsection*{\textnormal{\textit{Discontinuity in satellite overpass cycles.}}}

Overpass cycles for polar orbiting sensors are discontinuous. These sensors provide data for the vast majority of urban T\textsubscript{surf} study. Assessment of land T\textsubscript{surf} either sacrifices spatial resolution for a daily repeat cycle - as is the case with MODIS  - or sacrifices temporal resolution for a higher spatial resolution - ASTER or Landsat. These satellites cannot be used to assess the temporal development of sUHI without significant interpolation. Geostationary satellite sensors overcome this limitation but low spatial resolutions limit applicability in urban areas, especially at mid to high latitudes.

\subsection*{\textnormal{\textit{Clear-sky bias}}}

@@

As is the case with geometric biases, the magnitude of these temporal biases in remote sensed urban T\textsubscript{surf} is difficult to quantify without long-term ground truthing campaigns or significant interpolation. As such, the true temporal and geometric nature of urban effect on T\textsubscript{surf} and thus sUHI is presently unknown.

@@

\noindent In an attempt to answer these questions, this thesis

\begin{itemize}
	\item Develops and evaluates a method to retrieve atmospherically correction hemispherical radiometric urban surface temperatures from time-continuous measurements of upwelling longwave radiation.
	\item Compares urban surface temperatures and surface urban heat island magnitudes retrieved using the method to common remote sensed representations of the urban surface.
	\item Derives two long term climatology of hemispherical urban surface temperatures to observe seasonal and diurnal patterns of the surface urban heat island effect.
\end{itemize}
